---
title: 'Mitigating the risk of bias amplification for binary exposures when controlling for unmeasured spatial confounding'
author: "Danyang Li"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: cerulean
    code_folding: hide
    highlight: default
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
output_root <- "output/sim_results_mattern"
dir.create(output_root)
library(rlist)
library(magrittr)
library(tidyverse)
library(stringr)
library(sp)
library(geoR)
library(gridExtra)
library(fields)
R.files <- list.files("R/")
lapply(R.files, function(f){
  source(paste0("R/",f))
})
seed0<- 1121
```

## Simulation setup

### Data generating scenarios
```{r simsetup}
n_datasets <- 100
phi_seq <- c(0.04,0.15,0.6)
theta_seq <- c(0, 1, 3, 9)
params <- crossing(
    n_obs = n_datasets,
    phi_c = phi_seq, phi_u = phi_seq,
    theta = theta_seq,
    beta_true = 1,
    delta_c = sqrt(9/2),
    delta_u = sqrt(9/2)
)
params %>% mutate(Pns = theta^2/(theta^2+delta_c^2 + delta_u^2))
```

### Data generation (elemental processes before the assembly of X and Y)
```{r generate_data, eval=FALSE, echo=FALSE}
SAVE_RUNNING_INFO <- TRUE
library(parallel)
library(doParallel)
library(foreach)
library(doRNG)

message(output_root)
message(n_datasets)
message(Sys.time())

cl <- makeCluster(4)
registerDoParallel(cl)
output_dir <- paste0(output_root, "/datasets_new/")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
full_time0 <- Sys.time()
running_info <- tibble(n = 1024, full_time0, 
                       n_cluster = length(cl),
                       R_version = version %>% list())
if (SAVE_RUNNING_INFO) saveRDS(running_info, file = paste0(output_dir, "/runtime.rds"))
df_sim <- foreach(
    i = 1:n_datasets, .combine='bind_rows', .multicombine = TRUE, 
    .options.RNG = seed0,
    .inorder=FALSE,
    .verbose=FALSE,
    .errorhandling = "remove",
    .packages=c("magrittr", "tidyverse")) %dorng% {
        devtools::load_all()
      
        ## seeds
      seed_dataset <- 100*i
      set.seed(seed_dataset)

      df <- gen_data_comps(size = 1, n_obs = 1000, phi_seq = c(0.04, 0.15, 0.6),
                           sigsq_y_true_seq = 4, mat_corfun=mat_corfun_matern)
      out <- tidyr::crossing(dataset = i, seed_dataset, df)
      saveRDS(out, file = paste0(output_dir, i, ".rds"))
        
      out
    }
full_time1 <- Sys.time()
stopCluster(cl)
difftime(full_time1, full_time0)
running_info %<>% mutate(full_time1 = full_time1)
if (SAVE_RUNNING_INFO) saveRDS(running_info, file = paste0(output_dir, "/runtime.rds"))
```


### Data mapping
```{r map}
#df_comps <- gen_data_comps(n_obs = 1000, phi_seq=c(0.04,0.15,0.6), sigsq_y_true_seq = 4, mat_corfun = mat_corfun_matern)
dataset_id <- 1
df_comps <- readRDS(paste0(output_root,"/datasets_new/",
                           dataset_id,".rds"))
pred.grid <- expand.grid(seq(0,1,l=400),seq(0,1,l=400))
P = c(0, 0.5)
par(mfrow=c(2,2))
for (p in P) {
  df1 <- gen_data_from_comps(df_comps, beta=1, deltaE=NULL, coef_gen_y=1, 
                             coef_gen_c = sqrt(9/2), coef_gen_u=sqrt(9/2),
                             coef_inter=0,thresh=0, Pns=p, phi_c=0.6, phi_u=0.04,sigsq_y_true=4)
  df2 <- gen_data_from_comps(df_comps, beta=1, deltaE=NULL, coef_gen_y=1, 
                             coef_gen_c = sqrt(9/2), coef_gen_u=sqrt(9/2),
                             coef_inter=0,thresh=0, Pns=p, phi_c=0.04, phi_u=0.6,sigsq_y_true=4)
  if(p==0){
    expr = expression(paste(Pns==0,",",phi^c == 0.6,",", phi^u == 0.04))
    surface_kc(df1, "z_c", ini=c(1,0.6), pred.grid, expression = expr, 
               file.path = paste0(output_root,"/map/","data",dataset_id,"_phic",0.6,
                                  "_phiu",0.04,"_Pns",p,".rds"))
    expr = expression(paste(Pns==0,",",phi^c == 0.04,",", phi^u == 0.6))
    surface_kc(df2, "z_c", ini=c(1,0.04), pred.grid, expression = expr,
               file.path = paste0(output_root,"/map/","data",dataset_id,"_phic",0.04,
                                  "_phiu",0.6,"_Pns",p,".rds"))
  }else{
    expr = expression(paste(Pns==0.5,",",phi^c == 0.6,",", phi^u == 0.04))
    surface_kc(df1, "z_c", ini=c(1,0.6), pred.grid, expression = expr,
               file.path = paste0(output_root,"/map/","data",dataset_id,"_phic",0.6,
                                  "_phiu",0.04,"_Pns",p,".rds"))
    expr = expression(paste(Pns==0.5,",",phi^c == 0.04,",", phi^u == 0.6))
    surface_kc(df2, "z_c", ini=c(1,0.04), pred.grid, expression = expr,
               file.path = paste0(output_root,"/map/","data",dataset_id,"_phic",0.04,
                                  "_phiu",0.6,"_Pns",p,".rds"))
  }
}

```

### Model considered for comparison
```{r simmodel}
model_info <- bind_rows(
    tibble(method = "NS", class = "Non-spatial"),
    tibble(method = "true", class = "True-model"),
    tibble(method = "FDF", class = "Spatial spline", 
           Knots = c(5, 10, 25, 50, 100, 250, 500)),
    tibble(method = "PS", class = "Spatial spline", 
           Knots = c(500)),
    tibble(method = "EPS", class = "Spatial spline", 
           Knots = c(250, 500)),
    tibble(method = "1NN", class = "Matching", 
           Caliper = c(0, 0.08, 0.15, 0.6)),
    tibble(method = "Full", class = "Matching"),
    tibble(method = "1NN-PS", class = "Combined", 
           Knots = c(500), Caliper = c(0, 0.08, 0.15, 0.6)),
    tibble(method = "Full-PS", class = "Combined", 
           Knots = c(500))
)
model_info
```

## Implementation
### Iteration
```{r model_info, echo=FALSE}
model <- bind_rows(
  tibble(model = "Naive", inference = "freq"),
  tibble(model = "true_Model", inference = "freq"),
  tidyr::crossing(model = "est_df_y", inference = "freq",
                  n_knots = 500),
  tidyr::crossing(model = "est_df_x", inference = "freq",
                  n_knots = c(250, 500), caliper = NA, match = NA, clp = NA),
  tidyr::crossing(model = "fix_df", inference = 'freq', 
                  n_knots = c(5,15,25,50,100,250,500),  caliper = NA, match = NA, clp=NA),
  tibble(model = "Naive", inference = "freq",
         caliper = "distance", clp=c(0.08,0.15,0.6), match = "nearest"),
  tibble(model = "Naive", inference = "freq",
         caliper = "None", clp=99, match = "nearest"),
  tibble(model = "Naive", inference = "freq",
         caliper = "None", clp=99, match = "full"),
  tibble(model = "est_df_y", inference = 'freq',
         caliper = "None", clp=99, n_knots=NA, match = "nearest"),
  tibble(model = "est_df_y", inference = 'freq',
         caliper = "None", clp=99, n_knots=NA, match = "full")
)

iter_info <- tidyr::crossing(
  dataset = 1:n_datasets,
  phi_c = c(0.04,0.15,0.6),
  phi_u = c(0.04,0.15,0.6),
  P = c(0, 0.5, 0.9),
  coef_c = sqrt(9/2),
  coef_u = sqrt(9/2),
  model
) %>%
  mutate(iter = row_number(), 
         save_file= paste0(ifelse(!is.na(match), paste0(match,"_"),"_"),
                           ifelse(!is.na(caliper),paste0(caliper,clp,"_"),"_"),
                           model,"_",ifelse(!is.na(n_knots),n_knots,"_"))) %>%
  dplyr::select(iter, dataset, everything())
iter_info
```

### Model fitting
```{r model, eval=FALSE}
SAVE_ROUT <- FALSE
library(parallel)
library(doParallel)
library(foreach)
library(doRNG)

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
dir.create(output_root, recursive = TRUE, showWarnings = FALSE)
full_time0 <- Sys.time()
res <- foreach(
    i = 1:nrow(iter_info), .combine='bind_rows', .multicombine = TRUE, 
    .options.RNG = seed0,
    .inorder=FALSE,
    .verbose=FALSE,
    .errorhandling = "remove",
    .packages=c("magrittr", "tidyverse","lmtest",
                "mgcv","MatchIt","Rcpp","optmatch","sandwich")) %dorng% {
        devtools::load_all()
        suppressWarnings(rm(list = ls() %>% setdiff(c("iter_info", "i", "SAVE_ROUT", "output_root"))))
        gc()

        res0 <- iter_info[i, ]
        iter <- res0$iter
        stem <- iter
        
        output_dir <- paste0(output_root, "/",res0$save_file,"/")
        dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
        if (SAVE_ROUT) {
            sink(Rout_file)
        }
        
        ## Scenario
        cat("\n\nScenario:\n")
        glimpse(res0)
        
        ## Load in data ####
        cat("\n\nGenerating data:\n")
        df_comps <- readRDS(paste0(output_root,"/datasets_new/", res0$dataset, ".rds"))
        #ATE
        df <- res0 %$% gen_data_from_comps(df_comps, beta=1, deltaE=delta, coef_gen_y=1, 
                                           coef_gen_c = coef_c, coef_gen_u=coef_u,
                                           coef_inter=0,thresh=0, Pns=P, 
                                           phi_c=phi_c, phi_u=phi_u,
                                           sigsq_y_true=4)

        if_match <- !is.na(res0$match)
        time0 <- Sys.time()
        ## matched data ###
        if (if_match) {
          m <-  Matchit_list(df, caliper= res0$caliper, clp=res0$clp, 
                             rep = res0$match == "nearest", match_method = res0$match)
          if(res0$match == "nearest"){
            m <- cluster_nearest(m)
            m$df_dup$cluster_id <- m$df_dup$pair_id
          }
          df_ate <- m$df_dup
          ## weighted data ###
          df_w <- m$df_weight
          if(res0$match == "full"){
            df_w$cluster_id <- df_w$subclass
          }
        }

        ## fit model ####
        #seed_iter <- seed_dataset + iter
        cat("\n\nFitting model:\n")
        if(res0$model == "Naive"){
          df_orig <- df
          if(res0$match != "full"| is.na(res0$match)){
            if(if_match){df <- df_ate}
            #mod_indv <- lm(y_indv ~ X_indv, data = df)
            mod_lat <- lm(y_lat ~ X_indv, data = df)
            time1 <- Sys.time()
            #s_indv <- summary(mod_indv)
            s_lat <- summary(mod_lat)
            summ0 <- s_lat$coefficients["X_indv",] %>%
              set_names(c("ATE_est", "ATE_se", "indv_t", "indv_p")) %>% 
              as.list() %>%
              as_tibble() %>%
              mutate(param = "coef_x",
                     coef_se_rob = sqrt(sandwich::sandwich(mod_indv)[2, 2]),
                     bic = BIC(mod_indv), time0=time0, time1=time1)
            res_mod <- summ0
          }
        
          if(if_match) {
            if(res0$match == "full"){
              mod_indv_w <- lm(y_indv~X_indv, data = df_w, weights = df_w$weights)
              time1 <- Sys.time()
              time2 <- time3 <- time1
              s_w <- summary(mod_indv_w)
              D_se <- 0
              t_se <- 0
              res_mod <- c(time0=time0, time1=time1) %>% as.list() %>% as_tibble()
            } else {
              ## 1NN
              mp_diff <- df_orig$y_indv[m$index.treated] - df_orig$y_indv[m$index.control]
              n_match <- length(m$index.treated)
              D_se <- sqrt(sum((mp_diff - mean(mp_diff))^2)/((n_match-1)*n_match))
              D_se <- sd(mp_diff)
              trt_ut <- df_orig$y_indv[unique(m$index.treated)]
              crl_uc <- df_orig$y_indv[unique(m$index.control)] 
              
              w_trt=table(m$index.treated);w_crl=table(m$index.control)
              t_s2 <- var(trt_ut)*sum(as.numeric(w_trt)^2)/sum(as.numeric(w_trt))^2 +
                var(crl_uc)*sum(as.numeric(w_crl)^2)/sum(as.numeric(w_crl))^2
              t_se <-sqrt(t_s2)
            }
            
            res_mod <- c(D_se) %>% set_names("MPE")
          }
        } else if (res0$model == "true_Model") {
          mod_indv <- lm(y_indv ~ X_indv + z_c, data = df)
          time1 <- Sys.time()
          #s_indv <- summary(mod_indv)
          s_lat <- summary(mod_lat)
          summ0 <- s_lat$coefficients["X_indv",] %>%
              set_names(c("ATE_est", "ATE_se", "indv_t", "indv_p")) %>%
              as.list() %>% as_tibble() %>%
              mutate(param = "coef_x",
                     coef_se_rob = sqrt(sandwich::sandwich(mod_indv)[2, 2]),
                     bic = BIC(mod_indv))
            res_mod <- tibble(software = "lm") %>% crossing(summ0)
        } else if (res0$model == "fix_df") {
          if(if_match){
            df <- df_ate
            matchid_pair <- 1:length(m$index.treated)
            df$pair_id <- rep(matchid_pair,times=2)
          }
          k <- res0$n_knots + 1 ##for some reason 'k' knots translates to 'k-1' edf when looking at the "Approximate significance of smooth terms"
          #mod_indv <- gam(y_indv ~ X_indv + s(p0, p1, k = k, fx = TRUE), data = df)
          mod_lat <- gam(y_lat ~ X_indv + s(p0, p1, k = k, fx = TRUE), data = df)
          time1 <- Sys.time()
          summ0 <- extract_ests_gam(mod_lat, inter=FALSE, coef = "X_indv")
          res_mod <- summ0 %>% 
            mutate(param = "coef_X",software = "gam", time0 = time0, time1 = time1)
          if(if_match){                  
            time2 <- Sys.time()
            mod_indv_w <- gam(y_indv ~ X_indv + s(p0, p1, k = k, fx = TRUE), 
                              data = df_w, weights = df_w$weights_fq)
            time3 <- Sys.time()
            summ_w <- extract_ests_gam(mod_indv_w, inter=FALSE, coef = "X_indv")
            summ_w <- summ_w[-c(3,4)] %>% set_names(paste0("WLS_",colnames(summ_w)[-c(3,4)]))
            X =model.matrix.gam(mod_indv)
            df <- data.frame(cbind(df$y_indv,df$pair_id,X))
            colnames(df)[c(1,2)] <- c("y_indv","pair_id")
            mod_lm <- lm(y_indv ~ .-pair_id-X.Intercept., data = df)
            cluster_sem <- coeftest(mod_lm,vcov.=vcovCL,cluster=~pair_id)
            cluster_se <- cluster_sem["X_indv","Std. Error"]
            res_mod <- summ0 %>% bind_cols(summ_w) %>%
              mutate(ATE_cluster_se = cluster_se,
                     param = "coef_X",software = "gam", 
                     time0 = time0, time1 = time1,
                     time2 = time2, time3 = time3)
          }
        } else if (res0$model == "est_df_x") {
          k <- res0$n_knots
          #k <- ifelse(res0$n > 2000, 1000, res0$n/2)
          time0 <- Sys.time()
          mod_indv_x <-  gam(X_indv ~ s(p0, p1, k = k), data = df, family=binomial)
          
          timex <- Sys.time()
          mod_indv <-  gam(y_lat ~ X_indv + s(p0, p1, k = k), data = df,
                           sp = mod_indv_x$sp)
          time1 <- Sys.time()
          summ0 <- extract_ests_gam(mod_indv, inter=FALSE, coef = "X_indv")
      
          res_mod <- summ0 %>% 
            mutate(param = "coef_X",software = "gam", time0 = time0, timex = timex, time1 = time1)
          #res_mod <- data.frame(k=k, time0=time0, timex=timex, time1=time1)
          
        } else if (res0$model == "est_df_x_latent") {
          k <- res0$n_knots
          time0 <- Sys.time()
          mod_indv_x <-  gam(x_latent ~ s(p0, p1, k = k), data = df)
          timex <- Sys.time()
          mod_indv <-  gam(y_indv ~ X_indv + s(p0, p1, k = k), data = df,
                           sp = mod_indv_x$sp)
          time1 <- Sys.time()
          summ0 <- extract_ests_gam(mod_indv, inter=FALSE, coef = "X_indv")
      
          res_mod <- summ0 %>% 
            mutate(param = "coef_X",software = "gam", time0 = time0, timex = timex, time1 = time1)
        } else if (res0$model == "est_df_y") {
          cluster_ind = FALSE
          #res0$match != "full"
          if(res0$match != "full"| is.na(res0$match)){
            if(if_match){df <- df_ate; cluster_ind <- TRUE}
            #k <- length(unique(df$p0))/2
            k <- res0$n_knots
            mod_lat <- gam(y_lat ~ X_indv + s(p0, p1, k = round(k,0)), 
                            data = df)
            time1 <- Sys.time()
            summ0 <- extract_ests_gam(mod_lat, inter=FALSE, coef = "X_indv", 
                                      cluster = cluster_ind, match_data = df)
            res_mod <- summ0 %>% 
            mutate(param = "coef_X",software = "gam", 
                   time0 = time0, time1 = time1)
          }
          
          if(if_match){
            time2 <- Sys.time()
            k <- dim(df_w)[1]/2
            mod_indv_w <- gam(y_lat ~ X_indv + s(p0, p1, k = round(k,0)), 
                              data = df_w, weights = df_w$weights)
            time3 <- Sys.time()
            if(res0$match == "full"){
              time1 <- time3
              time2 <- time3 <- time1
              res_mod <- c("coef_X","gam") %>% 
                set_names(c("param","software"))%>% 
                as.list() %>% as_tibble() %>% mutate(time0 = time0, time1 = time1)
              cluster_ind <- TRUE
            }
            
            summ_w <- extract_ests_gam(mod_indv_w, inter=FALSE, coef = "X_indv",
                                       cluster = cluster_ind, match_data = df_w)
            summ_w <- summ_w[-c(3,4)] %>%
              set_names(paste0("WLS_",colnames(summ_w)[-c(3,4)]))
            #cluster_se_w <- coeftest(mod_indv_w, 
             #                        vcov. = vcovCL, cluster = ~subclass)
            res_mod %<>% bind_cols(summ_w) %>%
              mutate(time2 = time2, time3 = time3,k=k)
          }
        } else {
          res_mod <- tibble(message = "method not available")
        }
         
        ## Save and finish ####
        cat("\n\nSaving relevant output:\n")
        res_iter <- res0 %>%
            bind_cols(res_mod)
        saveRDS(res_iter, file = paste0(output_dir, 
                                        res0$model, stem, ".rds"))
        
        #cat("\n\nDone!\n")
        #glimpse(res_iter)
        ## return
        if (SAVE_ROUT) {
            #sink(type = "message")
            sink()
        }
        res_iter
    }
full_time1 <- Sys.time()
stopCluster(cl)
difftime(full_time1, full_time0)
```

### Model reults summary
```{r summary1, echo=FALSE, warning=FALSE, message = FALSE}
res_dir <- paste0(output_root,"/res_summary")
file_list <- list.files(path=res_dir)
replace = c("_est_df_x_"="EPS","_est_df_y_"="PS",
            "None99"="_","_fix_df_"="FDF",
            "_true_Model__"="true","_Naive__"="NS",
            "__partialtrue_Model__"="partial_true",
            "_Diff_AvgXLat__"="Diff_AvgXLat")
model_name = str_remove(str_replace_all(file_list, replace),
                        ".rds")
model_name = str_remove(str_remove(model_name,"distance"),"_")
for (i in seq_along(file_list)) {
  x <- readRDS(paste0(res_dir,"/",file_list[i]))
  assign(model_name[i],x)
}
```

```{r summary, results = "hide", warning=FALSE, message = FALSE}
res_list_constant <- list(NS, true, FDF5, FDF15, FDF25, FDF50, FDF100, FDF250, FDF500, PS500, EPS500, nearest_NS, nearest_PS_)

model_list <- c("Naive","true_Model",rep("fdf",7),"PS", rep("EPS",1),"M1",rep("M1_comb",1))#, "M1_weight")
phi_seq = c(0.04,0.15,0.6)
theta_seq = c(0,3,9)
model_name <- c("NS","true",paste0("FDF-",c(5,15,25,50,100,250,500)),"PS-500", paste0("EPS-", c(500)),
                "1NN","1NN-PS-500")

stat <- stat_all(res_list_constant, model_list, phi_seq, phi_seq,
                           theta_seq, ATE=1, rob_need=F)
result <- biasRMSE(stat, model_name, phi_seq,
                   phi_seq, theta_seq)
result$model <- factor(result$model,levels = model_name) 
result %<>% mutate(phic_expr = paste0("phi^c == ", phic),
                   phiu_expr = paste0("phi^u == ", phiu)) 

stat_se <- stat
se_name <- c("ATE_se_emp","ATE_se_est_mean",
             "ATE_se_MP_mean",
             "ATE_se_TS_mean",#
             "ATE_cluster_se_mean")
data_se = est_theta_se(stat_se, se_name, model_name, phi_seq, phi_seq, theta=c(0,3,9))
data_se$model <- factor(data_se$model,
                        levels = model_name)

data_se$theta <- factor(data_se$theta, levels = c("0","3","9"))
data_se_tb <-data_se
data_se_tb[,5:8] <- data_se_tb[,5:8]/data_se_tb$ATE_se_emp
data_se_tb %<>% dplyr::select(-c("ATE_se_emp"))
data_se_long <- data_se_tb %>% pivot_longer(cols = starts_with("ATE"),
                                            names_to = "est_method", values_to = "Value")
data_se_long$est_method=str_replace_all(data_se_long$est_method,c("ATE_se_est_mean"="LSE","ATE_se_MP_mean"="Matched-Pair",
                                          "ATE_se_TS_mean"="Two-Sample","ATE_cluster_se_mean"="Cluster-Robust"))

data_se_long %<>% mutate(phic_expr = paste0("phi^c == ", phi_c),
                         phiu_expr = paste0("phi^u == ", phi_u))

model_name1 = c("PS-500", "EPS-500", "1NN-PS-500")
modells <- list()
modells[[1]] <- PS500; modells[[2]] <- EPS500; modells[[3]] <- nearest_PS_ 
rate_df <- lapply(1:length(model_name1), function(l){
  res <- modells[[l]]
  res$lb <- res$ATE_est-1.96*res$ATE_se
  res$ub <- res$ATE_est+1.96*res$ATE_se
  res$cover <- (res$lb <= 1) & (res$ub >= 1)
  res <- res %>% group_by(theta, phi_c, phi_u) %>% summarise(cover_rate = mean(cover))
  res$model = model_name1[l]
  return(res)
})
rate_df <- do.call(rbind, rate_df)
```

## Visualization
### Bias and Variability

```{r bias, fig.height = 8, fig.width = 11}
pch = rep(c("NS/true","NS/true",rep("FDF",7),"PS",
            rep("EPS",1),"1NN",rep("1NN-PS",1)), times = 27)

method <- as.factor(pch)
pd <- position_dodge(0.5)
result$true = rep(result$mean[result$model=="true"], each = 13)

ggplot(result, aes(x=model, y= mean, colour = theta)) +
  geom_point(aes(shape=method),position = pd) + 
  geom_linerange(aes(ymin=lb, ymax=ub, group = theta),position = pd) +
  geom_hline(aes(yintercept=NS, color=theta), linetype=2) +
  geom_hline(aes(yintercept=true, color=theta), linetype="solid", color = "gray")+
  geom_hline(yintercept=1, linetype="solid", color = "gray") +
    theme_bw()+
    labs(y="Estimate of Exposure Effect", x = "Model")+
    theme(axis.text.x = element_text(angle = 90, vjust=0.4, hjust=0.95, size = 10),
        axis.ticks = element_blank(),
        strip.text = element_text(size = 15)) +
    facet_wrap( ~ phic_expr+phiu_expr,  labeller = labeller(.cols = label_parsed, .multi_line = FALSE))
```

### rMSE
```{r rMSE, fig.height = 8, fig.width = 11}
result2 <- result %>% filter(model != "true")
pch = rep(c("NS",rep("FDF",7),
            "PS",rep("EPS",1), "1NN", "1NN-PS"),  27)
method <- as.factor(pch)
ggplot(result2, aes(x=model, y= rMSE, colour = theta, group = theta)) +
  geom_point(aes(shape=method)) + 
  geom_line() +
  theme_bw() +
  labs(y="rMSE", x = "Model")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.4, hjust=0.95),
        axis.ticks = element_blank()) +
  facet_wrap( ~ phic_expr+phiu_expr,  labeller = labeller(.cols = label_parsed, .multi_line = FALSE))
```

### Uncertanity

i) SE est/SE empirical
```{r SE_ratio, fig.height = 8, fig.width = 11,warning=FALSE,}
data_se_long_select <- data_se_long %>%
  filter(model != "true", theta %in% c("0","3"))

data_se_long_select %>% 
  ggplot(aes(x = model, y=Value, color = est_method))+
  geom_point(aes(shape = theta))+
  geom_hline(yintercept = 1) +
  theme_bw() +
  labs(y="Uncertainty Estimate/Empirical Ratios", x = "Model")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.4, hjust = 0.95),
        axis.ticks = element_blank()) +
   facet_wrap( ~ phic_expr+phiu_expr,  labeller = labeller(.cols = label_parsed, .multi_line = FALSE))
```

ii) Coverage rate
```{r Coverage_rate,fig.height = 8, fig.width = 11}
rate_df %<>% mutate(phic_expr = paste0("phi^c == ", phi_c),
                   phiu_expr = paste0("phi^u == ", phi_u)) 
rate_df$model = factor(rate_df$model, levels = model_name)
rate_df$theta = factor(rate_df$theta, levels = c(0,1,3,9))
ggplot(rate_df%>% filter(theta != 1),aes(x = model, y = cover_rate, group=theta, color=theta))+
    geom_point()+
  #geom_line()+
  geom_hline(yintercept = 0.95, linetype = 2, color= "red")+
  theme_bw() +
  labs(y="Coverage Rate", x = "Model")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.4, hjust = 0.95),
        axis.ticks = element_blank())+
    facet_wrap( ~ phic_expr+phiu_expr,  labeller = labeller(.cols = label_parsed, .multi_line = FALSE)) 
```

## Additional simulation
### Sensitivity analysis (changing cutoff for latent exposure)
Spatial processes generation procedure and model fitting for this additional data scenarios are the same as before, so we just show the assembly of data processes

```{r Data_1, eval=FALSE}
SAVE_ROUT <- FALSE
library(parallel)
library(doParallel)
library(foreach)
library(doRNG)

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
dir.create(output_root, recursive = TRUE, showWarnings = FALSE)
full_time0 <- Sys.time()
res <- foreach(
    i = 1:nrow(iter_info), .combine='bind_rows', .multicombine = TRUE, 
    .options.RNG = seed0,
    .inorder=FALSE,
    .verbose=FALSE,
    .errorhandling = "remove",
    .packages=c("magrittr", "tidyverse","lmtest",
                "mgcv","MatchIt","Rcpp","optmatch","sandwich")) %dorng% {
        devtools::load_all()
        suppressWarnings(rm(list = ls() %>% setdiff(c("iter_info", "i", "SAVE_ROUT", "output_root"))))
        gc()

        res0 <- iter_info[i, ]
        iter <- res0$iter
        stem <- iter
        
        output_dir <- paste0(output_root, "/",res0$save_file,"/")
        dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
        if (SAVE_ROUT) {
            sink(Rout_file)
        }
        
        ## Scenario
        cat("\n\nScenario:\n")
        glimpse(res0)
        
        ## Load in data ####
        cat("\n\nGenerating data:\n")
        df_comps <- readRDS(paste0(output_root,"/datasets_new/", res0$dataset, ".rds"))
        #ATE
        df0 <- res0 %$% gen_data_from_comps(df_comps, beta=1, coef_gen_y=1, 
                                           coef_gen_c = coef_c, coef_gen_u=coef_u,
                                           coef_inter=0,thresh=0, Pns=P, 
                                           phi_c=phi_c, phi_u=phi_u,
                                           sigsq_y_true=4)
        new_cut <- quantile(df0$x_latent, 0.8)
        df <- res0 %$% gen_data_from_comps(df_comps, beta=1, coef_gen_y=1, 
                                           coef_gen_c = coef_c, coef_gen_u=coef_u,
                                           coef_inter=0,thresh=new_cut, Pns=P, 
                                           phi_c=phi_c, phi_u=phi_u,
                                           sigsq_y_true=4)
        

        ## Model fitting part is the same as before
         
        ## Save and finish ####
        cat("\n\nSaving relevant output:\n")
        res_iter <- res0 %>%
            bind_cols(res_mod)
        saveRDS(res_iter, file = paste0(output_dir, 
                                        res0$model, stem, ".rds"))
        
        #cat("\n\nDone!\n")
        #glimpse(res_iter)
        ## return
        if (SAVE_ROUT) {
            #sink(type = "message")
            sink()
        }
        res_iter
    }
full_time1 <- Sys.time()
stopCluster(cl)
difftime(full_time1, full_time0)
```

### Unmeasured spatial process in Y (not confounder)

```{r gen_data_app2, eval=FALSE}
SAVE_RUNNING_INFO <- TRUE
library(parallel)
library(doParallel)
library(foreach)
library(doRNG)

message(output_root)
message(n_datasets)
message(Sys.time())

cl <- makeCluster(4)
registerDoParallel(cl)
output_dir <- paste0(output_root, "/datasets_Zy/")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
full_time0 <- Sys.time()
running_info <- tibble(n = 1024, full_time0, 
                       n_cluster = length(cl),
                       R_version = version %>% list())
if (SAVE_RUNNING_INFO) saveRDS(running_info, file = paste0(output_dir, "/runtime.rds"))
df_sim <- foreach(
    i = 1:n_datasets, .combine='bind_rows', .multicombine = TRUE, 
    .options.RNG = seed0,
    .inorder=FALSE,
    .verbose=FALSE,
    .errorhandling = "remove",
    .packages=c("magrittr", "tidyverse")) %dorng% {
        devtools::load_all()
      
        ## seeds
      seed_dataset <- 100*i
      set.seed(seed_dataset)
      df <- gen_data_comps_phiY(size = 1, n_obs = 1000, phi_u=0.15, phi_c=0.6, phi_y=0.04,
                            sigsq_y_true_seq = 4, mat_corfun=mat_corfun_matern)
      out <- tidyr::crossing(dataset = i, seed_dataset, df)
      saveRDS(out, file = paste0(output_dir, i, ".rds"))
        
      out
    }
full_time1 <- Sys.time()
stopCluster(cl)
difftime(full_time1, full_time0)
running_info %<>% mutate(full_time1 = full_time1)
if (SAVE_RUNNING_INFO) saveRDS(running_info, file = paste0(output_dir, "/runtime.rds"))
```

```{r model_2, eval=FALSE}
SAVE_ROUT <- FALSE
library(parallel)
library(doParallel)
library(foreach)
library(doRNG)

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
dir.create(output_root, recursive = TRUE, showWarnings = FALSE)
full_time0 <- Sys.time()
res <- foreach(
    i = 1:nrow(iter_info), .combine='bind_rows', .multicombine = TRUE, 
    .options.RNG = seed0,
    .inorder=FALSE,
    .verbose=FALSE,
    .errorhandling = "remove",
    .packages=c("magrittr", "tidyverse","lmtest",
                "mgcv","MatchIt","Rcpp","optmatch","sandwich")) %dorng% {
        devtools::load_all()
        suppressWarnings(rm(list = ls() %>% setdiff(c("iter_info", "i", "SAVE_ROUT", "output_root"))))
        gc()

        res0 <- iter_info[i, ]
        iter <- res0$iter
        stem <- iter
        
        output_dir <- paste0(output_root, "/",res0$save_file,"/")
        dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
        if (SAVE_ROUT) {
            sink(Rout_file)
        }
        
        ## Scenario
        cat("\n\nScenario:\n")
        glimpse(res0)
        
        ## Load in data ####
        cat("\n\nGenerating data:\n")
        df_comps <- readRDS(paste0(output_root,"/datasets_Zy/", res0$dataset, ".rds"))
        #ATE
        df <- res0 %$% gen_data_from_comps_Zy(df_comps, beta=1, coef_y_c=1,
                                              coef_x_c = coef_c, coef_x_u=coef_u, coef_y_u=1,
                                              coef_inter=0, thresh=0, Pns=P,
                                              phi_c=phi_c, phi_u=phi_u, phi_y=phi_y,
                                              sigsq_y_true=4)

        ## Model fitting part is the same as before
         
        ## Save and finish ####
        cat("\n\nSaving relevant output:\n")
        res_iter <- res0 %>%
            bind_cols(res_mod)
        saveRDS(res_iter, file = paste0(output_dir, 
                                        res0$model, stem, ".rds"))
        
        #cat("\n\nDone!\n")
        #glimpse(res_iter)
        ## return
        if (SAVE_ROUT) {
            #sink(type = "message")
            sink()
        }
        res_iter
    }
full_time1 <- Sys.time()
stopCluster(cl)
difftime(full_time1, full_time0)
```

## Evaluation of correlation between fitted spline term and Zu/Zc

```{r IV, eval=FALSE}
SAVE_ROUT <- FALSE
library(parallel)
library(doParallel)
library(foreach)
library(doRNG)

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
dir.create(output_root, recursive = TRUE, showWarnings = FALSE)
full_time0 <- Sys.time()
res <- foreach(
    i = 1:nrow(iter_info), .combine='bind_rows', .multicombine = TRUE, 
    .options.RNG = seed0,
    .inorder=FALSE,
    .verbose=FALSE,
    .errorhandling = "remove",
    .packages=c("magrittr", "tidyverse","lmtest",
                "mgcv","MatchIt","Rcpp","optmatch","sandwich")) %dorng% {
        devtools::load_all()
        suppressWarnings(rm(list = ls() %>% setdiff(c("iter_info", "i", "SAVE_ROUT", "output_root"))))
        gc()

        res0 <- iter_info[i, ]
        iter <- res0$iter
        stem <- iter
        
        output_dir <- paste0(output_root, "/",res0$save_file,"/")
        dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
        if (SAVE_ROUT) {
            sink(Rout_file)
        }
        
        ## Scenario
        cat("\n\nScenario:\n")
        glimpse(res0)
        
        ## Load in data ####
        cat("\n\nGenerating data:\n")
        df_comps <- readRDS(paste0(output_root,"/datasets_new/", res0$dataset, ".rds"))
        #ATE
        df <- res0 %$% gen_data_from_comps(df_comps, beta=1, coef_gen_y=1, 
                                           coef_gen_c = coef_c, coef_gen_u=coef_u,
                                           coef_inter=0,thresh=0, Pns=P, 
                                           phi_c=phi_c, phi_u=phi_u,
                                           sigsq_y_true=4)
        

        cat("\n\nFitting model:\n")
        if (res0$model == "fix_df") {
          k <- res0$n_knots + 1
          mod_indv <- gam(y_indv ~ X_indv + s(p0, p1, k = k, fx = TRUE), data = df)
          
          Tmat <- predict(mod_indv, type = "terms", se.fit = FALSE)
          s_col <- grep("^s\\(p0,\\s*p1\\)$", colnames(Tmat))
          s_hat <- drop(Tmat[, s_col])
          cor_s_Zc <- cor(s_hat, df$z_c)
          cor_s_Zu <- cor(s_hat, df$z_u)
          
          summ0 <- extract_ests_gam(mod_indv, inter=FALSE, coef = "X_indv") %>% dplyr::select(c("ATE_est"))
          summ0$cor_Zc <- cor_s_Zc; summ0$cor_Zu <- cor_s_Zu
          res_mod <- summ0
        } else {
          res_mod <- tibble(message = "method not available")
        }
         
        ## Save and finish ####
        cat("\n\nSaving relevant output:\n")
        res_iter <- res0 %>%
            bind_cols(res_mod)
        saveRDS(res_iter, file = paste0(output_dir, 
                                        res0$model, stem, ".rds"))
        
        #cat("\n\nDone!\n")
        #glimpse(res_iter)
        ## return
        if (SAVE_ROUT) {
            #sink(type = "message")
            sink()
        }
        res_iter
    }
full_time1 <- Sys.time()
stopCluster(cl)
difftime(full_time1, full_time0)
```
